---
title: "Merging data for weighted metrics"
author: "Joe Brown"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal 

Here we want to merge data to get a list of data frames with weighted metrics.

We want the resulting object to be a list of lists (1-219) where each element is a list of SSP scenario data frames containing metric values and weights associated with each. The weights were those found using different combinations of criteria weights. Therefore, the result should be a list (of 219) where each element is a list of SSP scenarios with the following columns: `run_number`, `metric_result`, and `mc_weight`. For each list (1-219) the weights were calculated using a different combination of criteria weights; which can be found in the `criteria_wts` data file.


# Load libraries, data, and source files

## Libraries
```{r load libraries}
library(tidyverse)
library(spatstat)
library(tidyverse)
library(matilda)
```

## Source helper functions 

```{r, source helper functions}
source("source/source_all.R")
```

## Load data

```{r load data}
weight_list <- readRDS("data/model_wts.RDS")
metric_list <- readRDS("data/warming_results.RDS")
criteria_weights <- readRDS("data/criteria_wts.RDS")

```

# Merge metrics and weights

For this step we need to merge `weight_list` with `metric_list` by the `run_number` column in the data frames. 

To do this we use the `merge_dfs()` function that we wrote in `helper_functions.R`.

```{r merge weights and metric values}
weighted_metric_list <- lapply(weight_list, function(sample) {
  
  merge_dfs(sample, metric_list)

  })

```

```{r, merge criteria combinations and add scenario column}
# Use Map to merge the criteria_weights with weighted_metric_list
weighted_scenario_metrics <- Map(function(element, criteria) {
  
  lapply(names(element), function(name) {
    
    # copy data
    df <- element[[name]]
    
    # Add the criteria weights as new columns
    df$temp_wt <- criteria$temp_wt
    df$co2_wt <- criteria$co2_wt
    df$ocean_uptake_wt <- criteria$ocean_uptake_wt
    
    # add scenario names column
    df$scenario <- name
    
    return(df)
  
    })
  
}, weighted_metric_list, criteria_weights)
```

```{r save the new combined data}
saveRDS(weighted_scenario_metrics, "data/weighted_scenario_metrics.RDS")
```

