---
title: "Computing Summary statistic and error terms"
author: "Joe Brown"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

First computing summary statistics for the temperature metrics and then using IPCC values to compute error terms for each criteria weight combination.

*Note: IPCC values referenced here are stored in the setup.R file* 

# Compute summary statistics

Use `split_data_frame` to compute summary statistics (weighted median and percentiles) for temperature metrics from each SSP scenario. 

```{r}
stat_summary <- lapply(scenario_metrics, function(df) {
  
  split_data <- split_data_frame(df)
  
  data_summary_test(split_data)

})
```

## Compute error terms

Needs notes and documentation.

```{r}
error_results <- lapply(names(stat_summary), function(scenario_name) {
  
  scenario <- stat_summary[[scenario_name]]  # Get metric stats for the scenario
  
  # Process each criteria influence
  lapply(names(scenario), function(criteria_id) {
    metric_stats <- scenario[[criteria_id]]  # Get the data frame for the criteria
    
    # Retrieve corresponding IPCC values for the 'long' timescale for this scenario
    ipcc_values_for_scenario <- ipcc_values[[scenario_name]]
    ipcc_median <- ipcc_values_for_scenario[1]
    ipcc_lower  <- ipcc_values_for_scenario[2]
    ipcc_upper  <- ipcc_values_for_scenario[3]
    
    # Calculate the errors for each statistic (median, lower, upper)
    median_error <- compute_error(metric_stats$median, ipcc_median)
    lower_error  <- compute_error(metric_stats$lower, ipcc_lower)
    upper_error  <- compute_error(metric_stats$upper, ipcc_upper)
    
    # Return the error results for this criteria
    list(
      median_error = median_error,
      lower_error  = lower_error,
      upper_error  = upper_error
    )
  })
})

```

## Putting things together

Where I am at now is needing to put thins together. Ideally I think the best way is a data frame with the following column:

1. `scenario` - name of the SSP scenario.
2. `temp_wt` - the weight values given to observed temperature criterion. 
3. `co2_wt` - the weight values given to observed co2 criterion.
4. `ocean_uptake_wt` - the weight values given to observed ocean C uptake criterion.
5. `metric_stat_values` - the median, lower, and upper bounds from the Hector simulations. Nested in stat_summary (not sure how immediately critical these values are - don't think they are needed for ternary plot).
6. `absolute_error` - the absolute error between the metric estimates from Hector/Matilda and IPCC values. Should be one for each criterion weight combination.
7. `percentage_erro` - represents the percent error rather than absolute error.
