---
title: "Computing Summary statistic and error terms"
author: "Joe Brown"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

First computing summary statistics for the temperature metrics and then using IPCC values to compute error terms for each criteria weight combination.

*Note: IPCC values referenced here are stored in the setup.R file* 

# Compute summary statistics

Use `split_data_frame` to compute summary statistics (weighted median and percentiles) for temperature metrics from each SSP scenario. 

```{r}
stat_summary <- lapply(scenario_metrics, function(df) {
  
  split_data <- split_data_frame(df)
  
  data_summary_test(split_data)

})
```

## Compute error terms

Needs notes and documentation.

```{r}
# Compute error results using lapply
error_results <- lapply(names(stat_summary), function(scenario_name) {
  
  scenario <- stat_summary[[scenario_name]]  # Get metrics for the scenario
  
  # Use lapply for criteria weight combinations
  error_by_criteria_id <- lapply(names(scenario), function(criteria_id) {
    criteria_data <- scenario[[criteria_id]]  # Get the data frame for the criteria

    # Use lapply for term lengths and calculate errors
    error_by_term <- lapply(names(criteria_data), function(term_length) {
      metric_stats <- criteria_data[[term_length]]  # Get the metric stats for the term length
      
      # Retrieve corresponding IPCC values for this scenario and term length
      ipcc_values_for_scenario <- ipcc_term_values[[scenario_name]][[term_length]]
      
      # Calculate the errors for each statistic (median, lower, upper)
      median_error <- compute_error(metric_stats$median, ipcc_values_for_scenario[1])
      lower_error  <- compute_error(metric_stats$lower, ipcc_values_for_scenario[2])
      upper_error  <- compute_error(metric_stats$upper, ipcc_values_for_scenario[3])
      
      # Return error results as a data frame
      data.frame(
        criteria_id = criteria_id,
        term_length = term_length,
        median_error = median_error$percentage_error,  # Adjust if you want to include all error types
        lower_error = lower_error$percentage_error,
        upper_error = upper_error$percentage_error
      )
    })
    
    # Combine results for each term length into one data frame
    bind_error <- do.call(rbind, error_by_term)
    
    return(bind_error)  # Return the combined data frame for the criteria
  })
  
  # Combine all criteria results for the scenario into one data frame
  bind_criteria_error <- do.call(rbind, error_by_criteria_id)

  return(bind_criteria_error)  # Return the combined data frame for the scenario
  
})


# Set top level names of list to match IPCC naming
names(error_results) <- ssp_names

```

## Putting things together

Where I am at now is needing to put thins together. Ideally I think the best way is a data frame with the following column:

1. `scenario` - name of the SSP scenario.
2. `temp_wt` - the weight values given to observed temperature criterion. 
3. `co2_wt` - the weight values given to observed co2 criterion.
4. `ocean_uptake_wt` - the weight values given to observed ocean C uptake criterion.
5. `metric_stat_values` - the median, lower, and upper bounds from the Hector simulations. Nested in stat_summary (not sure how immediately critical these values are - don't think they are needed for ternary plot).
6. `absolute_error` - the absolute error between the metric estimates from Hector/Matilda and IPCC values. Should be one for each criterion weight combination.
7. `percentage_error` - represents the percent error rather than absolute error.

Where are these pieces of data currently?:

1. `scenario` - in ssp_names
2. `temp_wt` - in `criteria_weights` 
3. `co2_wt` - in `criteria_weights` 
4. `ocean_uptake_wt` - in `criteria_weights` 
5. `metric_stat_values` - skipping this for now -- will need it for another figure later.
6. `absolute_error` - in `error_results`
7. `percentage_error` - in `error_results`


```{r}
# building data frame for each SSP to plot

plotting_list <- lapply(names(error_results), function(df_name) {
  
  data = error_results[[df_name]]
  
  data$temp_wt = rep(criteria_weights$temp_wt, each = 3)
  data$CO2_wt = rep(criteria_weights$co2_wt, each = 3)
  data$ocean_uptake_wt = rep(criteria_weights$ocean_uptake_wt, each = 3)

  return(data)  
})

# Set top level names of list to match IPCC naming
names(plotting_list) <- ssp_names
```

Save plotting data:
```{r}
# save plot data 
saveRDS(plotting_list, "data/plotting_list.rds")
```

