---
title: "Plotting Results"
author: "Joe Brown"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

Plot results in ternary plot that shows the error associated with different numerical combinations of criteria weights. 


# Plots 

Testing out ternary plot code:
```{r}
# Create the ternary plot with error as color
ggtern(data = subset(plotting_list$`SSP1-1.9`, term_length == "short"), aes(x = temp_wt, y = CO2_wt, z = ocean_uptake_wt, color = median_error)) +
  geom_point(size = 1, shape = 15) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_light() +
  theme_showarrows() +
  theme_arrowlarge() +
  theme_hidetitles() +
  xlab("Temperature Weight") +
  ylab("CO2 Weight") +
  zlab("Ocean C Uptake Weight") +
  labs(
    title = "Error of central value",
    color = "Percent Error"
  )

ggsave("example_figure.png", 
       device = "png", 
       height = 4.5, 
       width = 7.29, 
       units = "in")

```

```{r, results='hide'}
term_lengths_list <- list("short", "mid", "long")
error <- list("median_error", "lower_error", "upper_error")

lapply(names(plotting_list), function(scenario_names) {
  print(paste("processing scenario:", scenario_names))

  data = plotting_list[[scenario_names]]
    
  lapply(term_lengths_list, function(term_length) {
    print(paste("processing term length:", term_length))
    
    lapply(error, function(error_term) {
      print(paste("processing error term:", error_term))
    
      plot_ternary(data, term_length = term_length, error = data[[error_term]], 
               filepath = paste0("figures/", scenario_names, "_", term_length, "_", error_term, ".png"))
      
    })
  })
})
```


### Identifying best criterion weight values

What I am trying to do -- subset the data by rules:
rule 1: median values less than 5 (within the +/- 5% range of IPCC)
rule 2: upper and lower values less that 10 (within the +/- 10% range of IPCC)

```{r}
# Generate a list of data frames containing the lowest error row for each term_length and scenario
lowest_error_list <- lapply(plotting_list, function(scenario) {

  # Define the term lengths
  term_lengths <- c("short", "mid", "long")

  # Process each term_length individually
  term_results <- lapply(term_lengths, function(term) {
    
    # Filter rows for the specific term_length and where lower_error and upper_error are both less than 10
    filtered_data <- subset(scenario, term_length == term & lower_error < 30 & upper_error < 30 & median_error < 15)

    # Return the resulting row (or empty data frame) for this term_length
    return(filtered_data)
  })

  # Name each element in the list by its term length (short, mid, long)
  names(term_results) <- term_lengths
  
  # Return the list of data frames for each term_length in this scenario
  return(term_results)
})

```

df of best results for each scenario term length:
```{r}
term_lengths <- c("2021-2040", "2041-2060", "2061-2100")

closest_IPCC <- data.frame(
  "scenario" = c(rep(ssp_names, each = 3)),
  "term" = c(rep(term_lengths))
)

median_values <- c(stat_summary$`SSP1-1.9`$`0.3.0.0.7`$short$median, 
                   stat_summary$`SSP1-1.9`$`0.3.0.0.7`$mid$median,
                   stat_summary$`SSP1-1.9`$`0.4.0.0.6`$long$median,
                   stat_summary$`SSP1-2.6`$`0.35.0.15.0.5`$short$median,
                   stat_summary$`SSP1-2.6`$`0.25.0.0.75`$short$median
                   
                   )

lower_values <- c(stat_summary$`SSP1-1.9`$`0.3.0.0.7`$short$lower, 
                  stat_summary$`SSP1-1.9`$`0.3.0.0.7`$mid$lower, 
                  stat_summary$`SSP1-1.9`$`0.4.0.0.6`$long$lower,
                  stat_summary$`SSP1-2.6`$`0.35.0.15.0.5`$short$lower)

upper_values <- c(stat_summary$`SSP1-1.9`$`0.3.0.0.7`$short$upper, 
                  stat_summary$`SSP1-1.9`$`0.3.0.0.7`$mid$upper, 
                  stat_summary$`SSP1-1.9`$`0.4.0.0.6`$long$upper, 
                  stat_summary$`SSP1-2.6`$`0.35.0.15.0.5`$short$upper)

merge(closest_IPCC, )
```

