---
title: "Re-run Matilda"
author: "Joe Brown"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal


```{r}
# lowest error weight_list:

lowest_errorlapply(lowest)
```

Combine the best weights list to the initial params:
```{r}
weighted_params_list <- lapply(best_weight_result_list, function(scenario) {
  
  params$run_number <- 1:nrow(params)
  
  weighted_params <- merge(params, scenario, by = "run_number")
  
  return(weighted_params)
  
})
```

Select the top 100 params
```{r}
top_params <- lapply(weighted_params_list, function(weighted_params){
  
  ranked_params <- weighted_params[order(-weighted_params$mc_weight), ]
  
  top_100 <- head(ranked_params, 100)
  
  top_100 <- select(top_100, -mc_weight, -run_number)
  
  return(top_100)
})
```

Re-run Matilda using the newly identified "top" parameter sets

```{r}
params_list <- top_params

# Initialize cluster
cl <- makeCluster(detectCores() - 1)

# Export required functions and objects to the cluster
clusterExport(cl, c("params_list", "ssp_list", "newcore", "reset", "iterate_model"))

# Start timer
start_time <- Sys.time()

# Run the model for each parameter set
new_model_result <- parLapply(cl, names(params_list), function(param_name) {
  
  # Extract current parameter set and its SSP scenario
  params <- params_list[[param_name]]  # Parameter set
  ssp_name <- sub("_.*", "", param_name)  # Extract SSP from param_name (e.g., "SSP1-1.9")
  
  # Get the associated INI file for this SSP
  scenario <- ssp_list[[ssp_name]]
  
  # Initialize Hector core
  core <- newcore(scenario, name = param_name)  # Use param_name to ensure uniqueness
  
  # Split the parameter set into chunks
  param_chunks <- split(params, 1:50)  # Adjust chunking if necessary
  
  # Run the model using parameter chunks
  result_list <- lapply(param_chunks, function(chunk) {
    
    reset(core)
    
    # Run the model and save outputs
    iterate_model(core = core, 
                  params = chunk, 
                  save_years = 1800:2100, 
                  save_vars = c("gmst", "global_tas", "CO2_concentration", "ocean_uptake"))
  })
  
  # Ensure the correct run numbers are added to each parameter chunk
  for (i in 2:length(result_list)) {
    max_run_number <- max(result_list[[i - 1]]$run_number)
    result_list[[i]]$run_number <- result_list[[i]]$run_number + max_run_number
  }
  
  # Combine chunks into a single result for this parameter set
  result <- do.call(rbind, result_list)
  
  # Tag results with parameter set name
  result$param_set <- param_name
  
  return(result)
})

# Stop timer
runtime <- Sys.time() - start_time
print(runtime)

# Stop the cluster
stopCluster(cl)
```

# Normalize model results to 1995-2014 reference period

```{r calculate global tas relative to 1995-2014}
new_rel_model_results <- lapply(new_model_result, function(df) {
  
  # subset to only include global_tas
  data_subset <- subset(df, 
                        variable == "global_tas")
  
  # compute normalize data
  normalized_data <- normalize_to_reference(data_subset,
                                            var = "global_tas",
                                            ref_start = 1995,
                                            ref_end = 2014)
  
  # replace values in results with normalized data
  data_subset$value <- normalized_data
  
  return(data_subset)
})

```

We now have the `rel_model_results`, the `value` column in each data frame has been replaced with `global_tas` values relative to the specified reference period.

# Compute metrics 

Using the `warming_metric` and `rel_model_results` we compute the end-of-century warming metrics for each ensemble member.

```{r compute wamring metrics for each ensemble member}
new_warming_results <- lapply(new_rel_model_results, function(df) {
  
  # Extract the term (short, medium, long) from the scenario column
  df$term <- gsub(".*_(short|mid|long)$", "\\1", df$scenario)
  
  # Apply different metric calculation based on the extracted term
  if ("short" %in% df$term) {
    new_metrics <- metric_calc(df %>% filter(term == "short"), metric_list$short)
  } else if ("mid" %in% df$term) {
    new_metrics <- metric_calc(df %>% filter(term == "mid"), metric_list$mid)
  } else if ("long" %in% df$term) {
    new_metrics <- metric_calc(df %>% filter(term == "long"), metric_list$long)
  }
  
  return(new_metrics)
})
```

Compute median and CI to compare with IPCC:
```{r}
new_summary_stats <- lapply(new_warming_results, function(df) {
  
  summary_stats <- df %>% 
    summarize(
      median = quantile(metric_result, probs = 0.5), 
      lower = quantile(metric_result, probs = 0.05), 
      upper = quantile(metric_result, probs = 0.95)) %>% 
    ungroup()

  return (summary_stats)
  
})
```



## Test 

```{r}
test_params <- constrained_param_set$`1`[["SSP1-1.9"]]

test_params <- select(test_params, -mc_weight, -run_number)

```

```{r}
ini_test <- ssp_list$`SSP1-1.9` 
reset(core_test)
core_test <- newcore(ini_test, name = "SSP1-1.9")

test_result <- iterate_model(core = core_test, 
                             params = test_params, 
                             save_years = 1800:2100, 
                             save_vars = "global_tas")

test_result <- test_result %>% 
  mutate(
    value = normalize_to_reference(test_result, var = "global_tas", 1995, 2014))

#metrics 
long <- new_metric(var = "global_tas", years = 2081:2100, op = median)
metric_test <- metric_calc(test_result, metric = long)
metric_test <- merge(test_params, metric_test, by = "run_number")
metric_test_2 <- metric_calc(constrained_result, metric = short)
metric_test_2 <- merge(constrained_result, metric_test_2, by = "run_number")

quantile(metric_test_2$metric_result, probs = c(0.05, 0.5, 0.95))
quantile(metric_test$metric_result, probs = c(0.05, 0.5, 0.95))

```

```{r}
constrained_result <- merge(rel_model_results$`SSP1-1.9`, constrained_weights$`1`$`SSP1-1.9`, by = "run_number")




ggplot() +
  geom_line(data = test_result,
            aes(x = year, y = value, group = run_number)) +
  geom_line(data = constrained_result, 
            aes(x = year, y = value, group = run_number), 
            color = 'red')
```

